---
title: "water balance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Input}

run <- 'run3'
LCM_prop <- 0.35 # proportion of single-crop farmers 

wd <- getwd()
folder <- file.path(wd, run)
if(!exists(folder)){dir.create(folder)}
source(file.path("./", 'functions_v8.r'))

# Parameters & files

startDate <- '1984-10-01'
endDate <- '2017-09-30'
alloFactor <- 0.55
FDdy <- c("11-01", "02-01", "04-01")
crops_var <- c("LCB", "SCB", "LCM", "SCM")
crops_all <- c("LCB", "SCB", "LCM", "SCM", "ALF", "PCH")
total_area <- 51747#ha
total_area_third <- total_area/3
lead = 7

infolder <- file.path(wd, "input_files")
wrtFile <- file.path(infolder, "table_wr.txt")
reservoirInflow <- file.path(infolder, "9013_9047_accuvol.txt")
percentilesFile <- file.path(infolder, "percentiles_9013_9047_accuvol.txt")
SFforecastFile <- file.path(infolder, "SFprediction_ARIMA_v3.csv") 

startYear <- as.numeric(substr(startDate, 1, 4))
endYear <- as.numeric(substr(endDate, 1, 4))
hyears <- if(substr(startYear, 6, 7) %in% 1:9){hyear <- startYear:endYear}else{(startYear + 1):endYear}

```

```{r open_data}

# Open data

## open demand file (water requirements for each crop and year) and extract demands for all crops (values in mm)
wrt <- read.table(wrtFile, header = T)
cols <- which(colnames(wrt) %in% crops_all)
wrt <- wrt[, c(1, cols)]
wrt$dates <- as.Date(wrt$dates)
wrt$year <- as.numeric(substr(wrt$dates, 1, 4))

## total demand per year (per hectare)
wrt_year <- aggregate(wrt[,-c(1, ncol(wrt))], by = list(wrt$year), FUN = 'sum')

```


# 1. Water balance

### Perfect information
Determine for each of the years if there is enough water to plant maize or not. This is based on the single crop decision (preference LCM) in decision a and in the double crop decision (SCM after rainfed SCB)

1) Estimate total demand for variable crop (1/3 of the area, the other 2/3 are alfalfa and peach) 
  (a) LCM for single-crop farmers
  (b) SCM for double-crop farmers
2) Estimate availability
3) Compare maximum demand (1) and availability (2)

```{r}

tot_demand <- function(LCM_prop, SCM_prop){
  # This function calculates the total demand for the planted area [in hm3]
  
  # The variable crop surface (1/3 of total crop area) can be planted with LCM, SCM or rainfed crops. Maximum area planted with LCM or SCM depends on the proportion of each type of farmer.
  
  wrt_area <- cbind(year = wrt_year[,1], wrt_year[,-1] * total_area_third * 10^-5)
  wrt_area$LCM <- wrt_area$LCM * LCM_prop
  wrt_area$SCM <- wrt_area$SCM * SCM_prop

  wrt_area$totalD <- wrt_area$ALF + wrt_area$PCH + wrt_area$LCM + wrt_area$SCM
  return(wrt_area)
}

pinfo <- function(LCM_prop){
  
  # 1) TOTAL DEMAND
  
  SCM_prop <- 1 - LCM_prop
  wrt_area <- tot_demand(LCM_prop, SCM_prop)

  # 2) TOTAL AVAILABILITY

  # Availability at the end of the season (considering allocation factor)
  inflow <- read.table(reservoirInflow, header = T)
  A <- inflow[inflow$day==30 & inflow$month==9,]
  A$AnEff <- A$accu_vol * alloFactor


  # 3) COMPARISON
  DA <- merge(wrt_area, A[A$year %in% wrt_area$year,])
  DA$balance <- DA$AnEff - DA$totalD
  DA <- DA[which(DA$hyear == startYear + 1):which(DA$hyear == endYear),]
  
  # Save file
  outfile <- file.path(folder, paste0('output_WB_PI.txt'))
  write.table(DA, outfile)
  
  return(DA)
}

DA_pid <- pinfo(LCM_prop)

```


### Informed decisions (percentiles)

Balance at the end of the season based on percentiles
Perfect information of water requirements

```{r informed_reservoir}

resN <- function(DA){
  # Availability at decision dates 
  AIne <- availability(reservoirInflow, percentilesFile, startDate, endDate, write.file = T, outputReturn = "all")
  # Subset for the first days of the month
  AIne1 <- AIne[substr(AIne$date, 9, 10) == '01', ]
  # Extrapolated availability at the end of the season considering allocation factor
  AIne1$AnEff <- AIne1$AIne * alloFactor

  # Total demand (including ALF & PCH)
  Dtot <- DA[, c('year', 'totalD')]

  # Comparison
  resN <- compareAD('AnEff', AIne1, Dtot)

  resN[, c('Nov', 'Feb', 'Apr', 'Sep')]

  return(resN)
} 

resN_d <- resN(DA_pid)

```


# Informed decisions (forecast)

Balance at the end of the season based on probabilistic forecast of inflow to the reservoir (7 month lead) and percentiles
Perfect information for demand.

Different percentiles of the forecast ensemble extracted, in addition to the ensemble mean, to consider possible over or underestimation

```{r}

resF <- function(DA, seq_percentiles = seq(0.1, 0.9, 0.1)){
  
  # availability based on forecast & historic percentiles (gives accumulated volume after lead time)
  AIneF <- availabilityF(reservoirInflow, percentilesFile, SFforecastFile, startDate, endDate, write.file = T, outputReturn = "AnF")
  AIneF_Eff <- data.frame(dates = AIneF[, 1], AIneF[, -1] * alloFactor)
  ensPercs <- t(apply(AIneF_Eff[, -1], 1, quantile, seq_percentiles))
  colnames(ensPercs) <- c(paste0('ens', seq_percentiles * 100))
  AIneF_Eff <- data.frame(AIneF_Eff, ensPercs)

  # Total demand (including ALF & PCH)
  Dtot <- DA[, c('year', 'totalD')]

  # Comparison
  listNames <- c('SFp', 'ensMeanSF', colnames(ensPercs))
  resF_full <- lapply(listNames, compareAD, AIneF_Eff, Dtot)

  resF <- lapply(resF_full, subset, select = c('Nov', 'Feb', 'Apr', 'May', 'Sep'))
  names(resF) <- listNames
  return(resF)

}

resF_d <- resF(DA_pid)

```

```{r}

# Save balance to file

res2File <- function(resFname, DA, resN, resF_list){
  
  resF <- resF_list[[resFname]]
  
  dat <- data.frame(PI_balance = DA$balance, N_Nov = resN[, 'Nov'], N_Feb = resN[, 'Feb'], N_Apr = resN[, 'Apr'], N_May = resN[, 'May'], F_Nov = c(resF[, 'Nov'], NA), F_Feb = c(resF[, 'Feb'], NA), F_Apr = c(resF[,'Apr'], NA), F_May = c(resF[, 'May'], NA))
  
  outfile <- file.path(folder, paste0('output_WB_', resFname, '.txt'))
  write.table(dat, outfile)}

lapply(names(resF_d), res2File, DA_pid, resN_d, resF_d)

```

```{r plot}

# Combine tables

plot_facets <- function(){
  
  ensfiles <- list.files(folder, pattern = paste0('output_WB_ens.[0-9]'), full = T)
  meanfile <- list.files(folder, pattern = paste0('output_WB_ensMeanSF'), full = T)
  files <- c(meanfile, ensfiles)


  ## ggplot
  library(ggplot2)
  library(reshape2)


  ### get PI and N columns and convert to long data
  datPN <- read.table(files[1])
  datPN <- datPN[, grep('N_|PI', colnames(datPN))]
  colnames(datPN) <- gsub('_Nov', 11, colnames(datPN))
  colnames(datPN) <- gsub('_Feb', 2, colnames(datPN))
  colnames(datPN) <- gsub('_Apr', 4, colnames(datPN))
  colnames(datPN) <- gsub('_May', 5, colnames(datPN))
  colnames(datPN) <- gsub('_balance', '', colnames(datPN))
  cnames <- colnames(datPN)
  ldatPN <- melt(as.matrix(datPN))
  colnames(ldatPN) <- c('year', 'information', 'balance')
  ldatPN$level <- ifelse(ldatPN$information == 'PI', 'PI', 'N')

  ### get the F columns and convert to long data
  to_long <- function(f){
    dat <- read.table(f)
    dat1 <- dat[, grep('F_', colnames(dat))]
    if(length(colnames(dat1)) == 3) {
      colnames(dat1) <- c('F11', 'F2', 'F4')
      ldat <- melt(as.matrix(dat1))
      colnames(ldat) <- c('year', 'information', 'balance')
    }
    if(length(colnames(dat1)) == 0){
      ldat <- data.frame(year = rownames(dat), information = rep('F5', length(dat1)), balance = dat1)
    }
    if(length(colnames(dat1)) == 4) {
      colnames(dat1) <- c('F11', 'F2', 'F4', 'F5')
      ldat <- melt(as.matrix(dat1))
      colnames(ldat) <- c('year', 'information', 'balance')
    }
    ldat$year <- factor(ldat$year)
    level <- gsub('.*WB_', '', f)
    level <- gsub('.txt', '', level)
    ldat$level <- level
    return(ldat)
  }

  ldatF <- lapply(files, to_long)
  ldatF <- do.call(rbind, ldatF)

  ### combine

  ldats <- rbind(ldatPN, ldatF)
  ldats$level <- factor(ldats$level, levels = c('PI', "N", unique(ldatF$level)))
  ldats$year <- factor(ldats$year, levels = rev(unique(ldats$year)))


  ### palette and breaks
  paletteLength <- 50
  paleta <- colorRampPalette(c("indianred1", "white", "darkblue"))(paletteLength)
  minScale <- signif(min(ldats$balance, na.rm = T) - 4, 2)
  maxScale <- signif(max(ldats$balance, na.rm = T) + 4, 2)
  myBreaks <- c(round(seq(minScale, 0, length.out = ceiling(paletteLength/2) + 1)), 
              round(seq(maxScale/paletteLength, maxScale, length.out=floor(paletteLength/2))))

  ### plot
  zp1 <- ggplot(ldats, aes(x = information, y = year, fill = balance)) +
  geom_tile() +
  scale_fill_gradientn(colours = paleta, breaks = myBreaks[c(2, 50)]) +
  scale_x_discrete(position = 'top') +
  theme_bw() +
  facet_grid(. ~ level, scales = 'free')

  print(zp1)
}

plot_facets()

```
